\input{include.tex}

\markright{Piotr Suwara\hfill Homological algebra II: November 5, 2014\hfill}

\begin{document}
	
	\begin{theorem}
		Let $\C$ be an abelian category. Then $\K(\C)$
		(also $\K^+, \K^-, \K^b$) with standard translation functor
		and distinguished triangles is triangulated.
	\end{theorem}
	
	\begin{remark}
		$\Cone(U)$ fits as $Z$ in TR1.
	\end{remark}
	
	\begin{definition}[cohomological functor]
		Assume $\C$ is triangulated, $\A$ is abelian.
		Let ${F :\C \to \A}$ be an additive functor. We call it 
		{\em cohomological}
		if for any distinguished triangle
		$X \xrightarrow u Y \xrightarrow v Z \xrightarrow w T(X)$
		we have an exact sequence\\
		$\ldots \to F(T^i(X)) \to F(T^i(Y)) \to F(T^i(Z)) 
		\to F(T^{i+1}(X)) \to \ldots$.
	\end{definition}
	
	\begin{definition}
		Let $\C$ be a~triangulated category,
		$S$ a~localizing class of morphisms in $\C$.
		We say that $S$ is {\em compatible} with triangulation if
		\begin{itemize}
			\item $s \in S \iff T(s) \in S$,
			\item in TR3, $f, g \in S \implies h \in S$ for any $h$.
		\end{itemize}
	\end{definition}
	
	\begin{theorem}
		Let $\C$ and $S$ be as above. On $\C[S^{-1}]$
		we can define
		\begin{itemize}
			\item $T_S: \C[S^{-1}] \to \C[S^{-1}]$, $T_S = T$
			on objects and morphisms,
			i.e. $T(X \xleftarrow{s} Z \xrightarrow{f} Y)
			=T(X) \xleftarrow{T(s)} T(Z) \xrightarrow{T(f)} T(Y)$.
			\item 
			$X \xrightarrow u Y \xrightarrow v Z \xrightarrow w T(X)$
			is distinguished in $\C[S^{-1}]$ if it is isomorphic to
			a~distinguished triangle coming from $\C$.
		\end{itemize}
		Then $\C[S^{-1}]$ with the structure defined above is triangulated.
	\end{theorem}
	
	\begin{corollary}
		Derived category of an abelian category inherits
		the triangulated structure from the homotopy
		category of complexes.
	\end{corollary}






\end{document}
 
 
